\documentclass{article}
\usepackage{graphicx}
\usepackage[margin=1.5cm]{geometry}
\usepackage{amsmath}

\begin{document}
\twocolumn

\title{Tuesday Warm Up, Unit 0: Foundations and Fundamentals}
\author{Prof. Jordan C. Hanson}
\maketitle

\section{Memory Bank}
\small
\begin{itemize}
\item \textbf{Convolution}: this is an operation that characterizes the response $h[n]$ of a linear system.
\begin{equation}
y[i] = h[n] * x[n] = \sum_{j=0}^{M-1}h[j]x[i-j] \label{eq:conv}
\end{equation}
In words, the output at sample $i$ is equal to the produce of the system response $h$ and the input signal $x$, summed over the proceeding $M$ samples (from $j=0$ to $j=M-1$).
\item \textbf{Discrete Delta Function}, $\delta[n]$: A standard impulse response that contains one non-zero sample.  It has the following property:
\begin{equation}
x[n] = \delta[n] * x[n] \label{eq:conv2}
\end{equation}
\item \textbf{Discrete Fourier Transform}, for a sampled, digitized signal $x_n$:
\begin{equation}
X_{\rm k} = \sum_{n=0}^{N-1} x_n e^{-2\pi j (k/N) n}
\end{equation}
\item In DFT analysis, we often need to know the $\Delta t$, time duration for samples, and the sampling rate, $f_{\rm s}$.  Note that $1/f_{\rm s} = \Delta t$.
\item For a sinusoid of frequency $f$ (Hz), the period is $T = 1/f$ (seconds).
\end{itemize}

\section{Unit Conversions for Frequency, Period, and Sampling Rate}

\begin{enumerate}
\item Convert the following:
\begin{itemize}
\item 0.0005 seconds to ms
\item $2.5\times 10^{-8}$ seconds to ns
\item $6.25\times 10^{-7}$ seconds to $\mu$s
\item 0.0000125 seconds to ms
\end{itemize}
\item Convert the following:
\begin{itemize}
\item $12500$ Hz to kHz
\item $2.5\times 10^{8}$ Hz to GHz
\item $6.25\times 10^{7}$ Hz to MHz
\item $2500$ Hz to kHz
\end{itemize}
\item Given the following \textit{sampling rates,} give the time duration of samples:
\begin{itemize}
\item 128 kHz
\item 64 kHz
\item 500 MHz
\item 50 MHz
\end{itemize}
\item Given the following \textit{time duration of samples,} give the sampling rates:
\begin{itemize}
\item 0.1 ms
\item 0.01 ms
\item 0.1 $\mu$s
\item 0.01 $\mu$s
\end{itemize}
\end{enumerate}

\section{The Discrete Fourier Transform}

\begin{enumerate}
\item Type \verb+help fft+ in an \verb+octave+ command window.  Read about the various ways to input data into this function that computes the ``fast Fourier transform'' of the data.
\item Write a brief \verb+octave+ script that defines a sampling rate, time samples, and a vector of data representing a unit impulse, $\delta[n]$.
\item Pass the data into the \verb+fft()+ function, and store the output in a variable, \verb+X+.
\item Keep only the first half of the data output, \verb+X = X(1:end/2)+.
\item Let $f_{\rm s}$ be the sampling rate.  Define frequencies as \\ \verb+f = linspace(0,fs/2,N)+, where $N$ is the current length of \verb+X+.
\item Multiply the vector \verb+X+ by $1/f_{\rm s} = \Delta t$, then plot the absolute value of \verb+X+ versus the frequencies \verb+f+. What result do you observe?
\item Repeat this process with a sine wave input.
\end{enumerate}

\end{document}
